Improving Item-Item Similarity Estimation during Collaborative Filtering

### 2014-03-25 paper/paper.tex
edited last two paragraphs of experiment section

\subsection*{Experiment}
Ideally, we would know {\em a priori} the true similarity for a set of item
pairs. We could then generate similarity estimates and readily determine the
true error in these estimates. But both $TrueSimRand$ and $TrueSimPref$ are
theoretical constructs and not metrics we have access to in the real world.
Fortunately, we can achieve good approximations of either $TrueSim$ given
sufficient data. We therefore define $GoldSim = Sim_N$ where $N$ is a number
large enough that $GoldSim \approx TrueSim$. With a measurable $GoldSim$ as our
goal, we can implement a simple predictivity experiment.

To obtain good $GoldSim$ values, we work with subsets consisting of item pairs
with at least either $N=40$ and $80$ common users. These choices are a
compromise between having enough data for statistically significant results and
having a sufficiently large $N$ that $GoldSim \approx TrueSim$. Note that even
with $N=80$ the average absolute error in $PearSim$ (i.e. $\left<\left| PearSim
- GoldSim \right|\right>$) is $\approx 0.1$ (see Fig.  \ref{fig:MSE_SimR}). So
  while $PearSim$ and $CosSim$ are relatively stable after this many common
users, there does remain a significant level of variance.

Our off-line experiment simulates the passing of time, applying ratings to items
in the order they actually occured. The similarity of item pairs is then
estimated each time a user has rated both items such that $N$ is incremented by
one. In this way, we can compare various estimators by comparing their outputs
to each other and to the known $GoldSim$ value for each pair. We proceed like
this for each pair in a test set for $n=1,...,N$ to get varying degrees of
sparsity. We then compute the average squared error of the estimate for each
$n$.

>>>>

\subsection*{Experiment}
Ideally, we would know {\em a priori} the true similarity for a set of item
pairs. We could then generate similarity estimates and readily determine the
true error in these estimates. But both $TrueSimRand$ and $TrueSimPref$ are
theoretical constructs and not metrics we have access to in the real world.
Fortunately, we can achieve good approximations of either $TrueSim$ given
sufficient data. We therefore define $GoldPearSim = PearSim_N$ and $GoldCosSim =
CosSim_N$ where $N$ is a large enough number of common users that $GoldSim 
\approx TrueSim$.\footnote{To calculate $GoldSim$ in practice, we use $Sim$ with
all rating data, as long as the number of common users is greater than or equal
to $N$} With a measurable $GoldSim$ as our goal, we can implement 
a simple predictivity experiment.

To obtain good $GoldSim$ values, we work only with the item pairs that have at 
least either $N=40$ and $80$ common users. These choices are a compromise 
between having enough data for statistically significant results and having 
a sufficiently large $N$ that $GoldSim \approx TrueSim$.

In order to test predicted $TrueSim$, we construct an off-line experiment to
simulate observing sparse amounts of rating data. Our experiment reenacts  
the passing of time. Starting with all ratings withheld, we apply ratings to 
items in the order they actually occured. The similarity of item pairs is then
estimated each time the number of common users $n$ is incremented. In this way,
we can quanitatively compare various estimators by calculating their mean
squared error when approximating $GoldSim$ across pairs of items. Since 
different estimators may perform better for different degrees of sparsity, 
we compute the mean squared error for $n=1,...,N$ separately.
